---
title: "Ethiopia Serology"
output:
  html_document: 
    toc: true
    toc_float: true
    code_folding: hide
---

```{r, include=FALSE}
library(tidyverse)
library(scales)
knitr::opts_chunk$set(message = "")
```

#Base Survey Files#

## EA file to get basic sampling scheme

```{r}
ea = readxl::read_excel('MIS 2015 Data/EA.xlsx')
# View(ea)

glimpse(ea)

range(ea$`EA Code`)
range(ea$`EA Name`)

```

## Census file to get GPS coordinates of points (such as households), the name of the head of the household and a comment

- filter to key variables: _id,   
- instanceID , meta-rowID, houseNumber , placeName ,   
- location_altitude, location_latitude, location_longitude   

```{r}
cen_raw  = read_csv("MIS 2015 Data/census_mobile.csv")  

# str(cen_raw)

cen = cen_raw %>%
    select(
        `_id`, houseNumber , placeName , `_savepoint_timestamp` , 
        sampleFrame , selected, valid, 
        location_altitude, location_latitude, location_longitude
    ) %>%
    mutate(
        date = lubridate::ymd( as.Date(`_savepoint_timestamp` ) ) ,
        altitude = as.numeric( location_altitude), 
        latitude = as.numeric( location_latitude), 
        longitude = as.numeric( location_longitude)
    )

glimpse(cen)

n_distinct(cen$houseNumber)
range(cen$placeName)
```

# Survey Weights

```{r}
class1 = readxl::read_excel(
    'MIS 2015 Data/Copy of weight for MIS-2015_final.xls',
    sheet = 'class1') %>%
    rename( EA = `84` ,
            ClusterID = `CLUSTER-ID` ,
            weight.hh = `HH design weight W` ,
            stratNumber = `Stratum number` )

class2 = readxl::read_excel(
    'MIS 2015 Data/Copy of weight for MIS-2015_final.xls',
    sheet = 'class2') %>%
    rename( ClusterID = X__1 ,
            weight.hh = `HH design weight W` )
# View(ea)

glimpse(class1)
glimpse(class2)

weights = bind_rows( class1, class2 )

# strata
# count( weights, stratNumber )
# count(weights, TOWN) # 83
# count(weights, REGION) # 10
# count(weights, `UR/RU`) # 2
# count(weights, `SUBCITY`) # 2
# count(weights, CLASS) # 2

```

- The 'Stratum number' variable has `r length( unique(weights$stratNumber ) )` levels and will be used as the strata variable for survey estimations

## Join weights with ea file EA Name variable

```{r}

# 1. Link ea with  complete wieghts dataset  so that we can be sure the link correctly
# matches the region, zone, district, etc.

 eaw = left_join( ea, weights, by = c('EA Code' = 'ClusterID' ))
 
 # all match
 # anti_join( ea, weights, by = c('EA Code' = 'ClusterID' ))
 # anti_join(  weights, ea, by = c( 'ClusterID' = 'EA Code'))

 # glimpse(eaw)
 
 # confirm names ok
 # sum(eaw$Region %in% eaw$REGION)
 # count( eaw[ !eaw$Region %in% eaw$REGION,] , Region, REGION ) #OK
 # 
 # sum(eaw$Zone %in% eaw$ZONE)
 # count( eaw[ !eaw$Zone %in% eaw$ZONE,] , Zone, ZONE ) #OK
 # 
 # sum(eaw$Woreda %in% eaw$WEREDA)
 # count( eaw[ !eaw$Woreda %in% eaw$WEREDA,] , Woreda, WEREDA ) #OK
 
 # select needed variables
 
# 2. Repeat link with only the needed variables from weights
 eaw = left_join( ea, 
                   weights %>% select('ClusterID', `Stratum number`, weight.hh ), 
                   by = c('EA Code' = 'ClusterID' )) %>%
     mutate( ea = as.character(`EA Code`) )
 
 glimpse(eaw)
```

## Household file

- selecting key fields: household_id, member_id, gender, age_in_months, age_in_years, sleep_under_net, rdt_result, blood_slide, dbs, barcode, ethnicity 

```{r}
# get cell names and specify types

h_raw = readxl::read_excel("MIS 2015 Data/household mobile.xlsx", 
    sheet = "Duplicates Removed" ,
    col_types = c("text")
) 

glimpse(h_raw)

# _id versus household_id
# n_distinct(h_raw$`_id`)
# n_distinct(h_raw$household_id)
# sum( h_raw$`_id` == h_raw$household_id, na.rm = T)

# names(h_raw)

# h = h_raw %>%
#     select( household_id, member_id, gender, age_in_months, age_in_years, sleep_under_net,
#             rdt_result, blood_slide, dbs, barcode, ethnicity ) %>%
#     mutate( age_in_years = as.integer(age_in_years) , 
#             age_in_months = as.integer(age_in_months) 
#             )

h_id = h_raw %>% 
    # select( `_id`, household_id , `_savepoint_timestamp`, 
    #         ea , selected , region , zone , district, 
    #         gps_location_latitude, gps_location_longitude, gps_location_altitude 
    #         ) %>%
    mutate(
        date = lubridate::ymd( as.Date(`_savepoint_timestamp` ) ) ,
        altitude = as.numeric( gps_location_altitude), 
        latitude = as.numeric( gps_location_latitude), 
        longitude = as.numeric( gps_location_longitude)
    )    %>%
    select( -gps_location_latitude, 
            -gps_location_longitude, 
            -gps_location_altitude
            )

glimpse( h_id )

bind_cols( 
    ea =  n_distinct( h_id$ea),
    id = n_distinct( h_id$household_id) ,
    region = n_distinct( h_id$region)
)
```


## Load household member file get basic description of raw data

- selecting key fields: household_id, member_id, gender, age_in_months, age_in_years, sleep_under_net, rdt_result, blood_slide, dbs, barcode, ethnicity 

```{r}
# get cell names and specify types

hm_raw = readxl::read_excel("MIS 2015 Data/household_member_mobile.xlsx", 
    sheet = "Duplicates Removed" ,
    col_types = c("text")
) 

# glimpse(hm_raw)

# n_distinct(hm_raw$`_id`)
# n_distinct(hm_raw$member_id)
# sum( hm_raw$`_id` == hm_raw$member_id, na.rm = T)

hm = hm_raw %>%
    # select( household_id, member_id,  gender, age_in_months, age_in_years,
    #         live_here, stay_here_last_night, sleep_under_net,
    #         rdt_result, blood_slide, dbs, barcode, ethnicity ) %>%
    mutate( age_in_years = as.integer(age_in_years) , 
            age_in_months = as.integer(age_in_months) 
            )


```


## Create variables for Pf and Po (other plasmodia)

RDT result is coded as 1 = pf, 2 = Pv/Po/Pm, 3 = pf/mixed, 4 = negaative, 5 = invalid.

- Created pf_rdt variable: 1<- if 1 or 3 ; 0<- if 2 or 4   
- Created po_rdt (other) variable: 1 <- if 2 or 3 ; 0 <- if 1 or 4

```{r}

hm = hm %>%
    mutate(
        # pf 
        # if RDT = 1 (pf) or 3 (pf/mixed), then pf pos (1); 
        # if RDT = 2 (Pv/Po/Pm) or 4 (negative), then pf neg (0), 
        # if RDT = 5 (invalid) or missing, then NA
        pf_rdt = ifelse( rdt_result %in% c("1", "3") , 1, 
                     ifelse( rdt_result %in% c("2", "4") , 0 , NA ) 
                     ) ,
        
        # po
        # if RDT = 2 (Pv/Po/Pm) or 3 (pf/mixed), then po pos (1); 
        # if RDT = 1 (pf) or 4 (negative), then p0 neg (0), 
        # if RDT = 5 (invalid) or missing, then NA
        po_rdt = ifelse( rdt_result %in% c("2", "3") , 1, 
                     ifelse( rdt_result %in% c("1", "4") , 0 , NA ) 
                     ) 
        
        
    )
```


## Join datasets

- There are `r comma( nrow(hm) )` household member records  
- There are `r comma( nrow(h_id) )` household records  

```{r}

# ec = inner_join( ea, cen_raw, by = c('EA Code' = 'placeName') )
# nrow(ea); nrow(cen_raw); nrow(ec)
# 
# ch = inner_join(cen_raw, h_raw, by = c('_id' = '_id'))
# nrow(h_raw); nrow(ch)

# Join with h file by household_id, or _id?  more matches with _id
# n_distinct(hm$household_id)
# anti_join(hm, h_raw, by = c( 'household_id' = '_id' ))
# anti_join(hm, h_raw, by = c( 'household_id' = 'household_id' ))


hhm = inner_join(h_id, hm, by = c('_id' = 'household_id')) %>%
    select( -`_id` )

# glimpse(hhm)

# Check ea codes
count( hhm, ea) # 555 + NA

sum( is.na( hhm$ea)) # 878 NA

count( hhm, region, zone, district ) #328
# nrow(hm); nrow(hhm)



```

- There are `r comma( nrow(hhm) )` linked household-household member records  

- Join weights with house-houseMember-serology files

```{r}

hhmw = inner_join( eaw, hhm, by = 'ea' ) 

# unmatched
# anti_join( eaw, hhm,  by = 'ea') 
# anti_join(  hhm, eaw,  by = 'ea' ) 

```

- All eaw records match with hhm record  
- `r nrow(anti_join(  hhm, eaw,  by = 'ea' ) )` hhm records did not match with eaw because the ea is missing from `r sum(is.na(hhm$ea))` hhm records.  

# Serology 

## Serology dataset

```{r}

MIS_2015_Serology_Results <- readxl::read_excel("MIS 2015 Data/MIS 2015_Serology_Results.xlsx")

# View(MIS_2015_Serology_Results)

 sero_file = 'MIS 2015 Data/MIS 2015_Serology_Results.xlsx'
 
 sero =  readxl::read_excel( sero_file , 
                             sheet = "FI - Bkgd" , 
                             col_types = c( rep("text",3), rep('numeric',14) ) 
                             ) %>% 
     select( 1:17 ) %>%
     filter( !is.na(Barcode) )
 
 glimpse(sero)
 # Hmisc::describe( sero )
 # nrow(sero)
 
 # check for duplicate serology barcodes
    serology_dublicate_barcodes = count(sero, Barcode) %>% filter(n>1) 
    write_csv(serology_dublicate_barcodes, 
              'MIS 2015 Data/serology_duplicate_barcodes.csv' )
     
 # filter hhmw to those with barcode
 hhmwb = hhmw %>% filter( !is.na(barcode) )
 # glimpse(hmb)
 
 shm =inner_join( hhmwb, sero, by = c( 'barcode' = 'Barcode' ) )
 # glimpse(shm)
 
  ashm = anti_join( sero, hhmwb, by = c( 'Barcode' = 'barcode' ) ) %>%
      select( Barcode )
  write_csv(ashm, 
              'MIS 2015 Data/unmatched_serology_barcodes.csv' )
  # glimpse( ashm )
  
```

- `r nrow(serology_dublicate_barcodes)` records in the serology file have duplicate barcodes.

- `r nrow(ashm)` records are in the serology file but do not have a matching barcode in the combined household-household member file.

```{r}
    hshm = inner_join(h_id, shm, by = c('_id' = 'household_id'))
    # nrow(h_id); nrow(shm); nrow(hshm)

    # check for duplicate serology barcodes
    count(sero, Barcode) %>% filter(n>1) 
    
    
    # check for duplicate barcodes in MIS
    hmb_dublicate_barcodes = hhmwb %>%
        group_by( barcode) %>%
        mutate( n = n() ) %>%
        filter( n > 1 ) 
    
    hmb_dublicate_barcodes
    write_csv(hmb_dublicate_barcodes, 
              'MIS 2015 Data/hmb_dublicate_barcodes.csv' )
    
    
    hshm_dublicate_barcodes = count(hshm, barcode) %>% filter(n>1) 
    hshm_dublicate_barcodes
    write_csv(hshm_dublicate_barcodes, 
              'MIS 2015 Data/hshm_dublicate_barcodes.csv' )
    

    # are any of the hmb duplicates in the hshm duplicates?
    sapply( hmb_dublicate_barcodes$barcode ,
            function(x) grepl(x, paste( hshm_dublicate_barcodes$barcode, collapse = ' ' ) )
    )
    
  write_csv(hshm, "MIS 2015 Data/hshm.csv" )
  
  # hshm = read_csv("MIS 2015 Data/hshm.csv")
```

- `r nrow(hshm_dublicate_barcodes)` records in the combined household-household member file have duplicate barcodes.


## Summary of Linkage with Household/Household members

- There are `r comma(nrow(hmb))` household member records with a barcode   
- There are `r comma(nrow(sero))` serology records with a barcode  

- Combined, there are `r comma(nrow(hshm))` linked household member records with serology data

- `r percent( nrow(hshm) / nrow(sero) )` of serology results were linked with an individual from the survey

- All of the records have linked household data

**Can we link more household member records linked with serology?**
Unfortunately no, because there are no variables other than barcode to use for linking.  The failures to link are probably due to data entry errors.  As noted, there are a number of duplicate barcodes.  Additionally, the length of the barcode is not consistent:

`r   print( bind_rows(table( nchar(sero$Barcode) ), table( nchar( hmb$barcode))) )` 

So, one form of error appears to be a dropped or added character, although this accounts for only some of the linking failures.  

*It is important to note that there is also a possibility that, just as data entry error causes some records not to link, it is equally probable that some records are linked incorrectly.*  

**The combined linked file is "hshm.csv"**

Here is a list of the variables and first few data elements:

```{r}
 glimpse(hshm)
```

## Serology Cut-offs (taken from ppt and put into excel)

**TODO- Apply cutpoints to data set...**

```{r}
 cutoffs = readxl::read_excel( 'MIS 2015 Data/cutoffs.xlsx' ) 
 glimpse( cutoffs )
 
 hshm.c = hshm %>%
     mutate(
         `PfAMA1(43)`
     )
 
```


**Summarise**

# Data Summary (Unweighted)

## Summarise RDT by Age Group

note: pf_rdt.n is the number with a pf_rdt result;  po_rdt.n is the number with a po_rdt (other plasmodia) result.

```{r}
hshm.age = hshm %>%
    mutate( age_group = cut( age_in_years ,
                             c(0, 5, 10, 15, 20, 30, 40, 50, Inf),
                             c('<6', '6-10', '11-15', '16-20', '21-30', '31-40', '41-50', '50+' ) ,
                             include.lowest = TRUE
                             ) 
    ) %>%
    group_by( age_group ) %>%
    summarise(
        n = n(), 
        pf_rdt.n = sum( pf_rdt %in% 0:1 ) ,
        po_rdt.n = sum( po_rdt %in% 0:1 ),
        pf_rdt = sprintf( '%1.1f%%', 100 * mean(pf_rdt, na.rm = T) ),
        po_rdt = sprintf( '%1.1f%%', 100 * mean(po_rdt, na.rm = T) )
    )

knitr::kable(hshm.age)        

```


## Summarise RDT by Altitude

```{r}
hshm.alt= hshm %>%
    mutate( altitude.cat = cut( altitude ,
                             c(0, 1000, 2000, Inf),
                             c('<=1000', '1001-2000', '>2000' ) ,
                             include.lowest = TRUE
                             ) 
    ) %>%
    group_by( altitude.cat ) %>%
    summarise(
        n = n(), 
        pf_rdt.n = sum( pf_rdt %in% 0:1 ) ,
        po_rdt.n = sum( po_rdt %in% 0:1 ),
        pf_rdt = sprintf( '%1.1f%%', 100 * mean(pf_rdt, na.rm = T) ),
        po_rdt = sprintf( '%1.1f%%', 100 * mean(po_rdt, na.rm = T) )
    )

knitr::kable(hshm.alt )       

```

# Data Summary (Weighted)

## Summarise RDT by Age Group

note: pf_rdt.n is the number with a pf_rdt result;  po_rdt.n is the number with a po_rdt (other plasmodia) result.

```{r}
hshm.age = hshm %>%
    mutate( age_group = cut( age_in_years ,
                             c(0, 5, 10, 15, 20, 30, 40, 50, Inf),
                             c('<6', '6-10', '11-15', '16-20', '21-30', '31-40', '41-50', '50+' ) ,
                             include.lowest = TRUE
                             ) 
    ) %>%
    group_by( age_group ) %>%
    summarise(
        n = n(), 
        pf_rdt.n = sum( pf_rdt %in% 0:1 ) ,
        po_rdt.n = sum( po_rdt %in% 0:1 ),
        pf_rdt = sprintf( '%1.1f%%', 100 * mean(pf_rdt, na.rm = T) ),
        po_rdt = sprintf( '%1.1f%%', 100 * mean(po_rdt, na.rm = T) )
    )

knitr::kable(hshm.age)        

```


## Summarise RDT by Altitude

```{r}
hshm.alt= hshm %>%
    mutate( altitude.cat = cut( altitude ,
                             c(0, 1000, 2000, Inf),
                             c('<=1000', '1001-2000', '>2000' ) ,
                             include.lowest = TRUE
                             ) 
    ) %>%
    group_by( altitude.cat ) %>%
    summarise(
        n = n(), 
        pf_rdt.n = sum( pf_rdt %in% 0:1 ) ,
        po_rdt.n = sum( po_rdt %in% 0:1 ),
        pf_rdt = sprintf( '%1.1f%%', 100 * mean(pf_rdt, na.rm = T) ),
        po_rdt = sprintf( '%1.1f%%', 100 * mean(po_rdt, na.rm = T) )
    )

knitr::kable(hshm.alt )       

```

## Summarise travel history

- Load Eth Geo Area for travel history

```{r}

eageo = readxl::read_excel('MIS 2015 Data/Eth Geo Area.xlsx')
# View(eageo)
glimpse(eageo)
```


# MAP of Households

```{r, warning=FALSE}
library(leaflet)

pal <- colorNumeric(
  palette = "Blues",
  domain = range( as.numeric(h_id$ea), na.rm = TRUE ) 
)

m = leaflet( data = h_id  ) %>% 
               addProviderTiles("OpenStreetMap.France", group = "OpenStreetMap") %>%
               addProviderTiles("Stamen.TonerLite", group = "Toner Lite") %>%
               addProviderTiles( "OpenMapSurfer.Roads", group = "Roads" ) %>%
               addLabelOnlyMarkers(  
                    data =  h_id , 
                    ~longitude, 
                    ~latitude,
                    label =  'x' #~as.character( ea )
                ) %>%
         addCircles(
                  # popup = popup_cluster(),
                  lng = ~longitude, lat = ~latitude,
                  radius = 20, #~Survey_Weight*6e3,
                  stroke = TRUE,
                  weight = 1,
                  color = 'black',
                  opacity = .7 ,
                  fill = TRUE ,
                  fillColor =  ~pal( as.numeric(ea) ),
                  fillOpacity = .7 ,
                  group = "Sampling clusters"
               ) %>%
               addLayersControl(
                  # overlayGroups = c("Admin areas", "Sampling clusters", "Data Labels", "Admin Labels"),
                  baseGroups = c("OpenStreetMap", "Toner Lite", "Roads"),
                  options = layersControlOptions(collapsed = FALSE),
                  position = "bottomleft"
               )  
# %>%    
#     addLegend(
#         position = 'bottomright',
#         pal = pal ,
#         values = ~as.numeric(ea)  , 
#         opacity = 1,
#         title = 'An Obvious Legend'
#     )

m


```

